---
title: "Poisson Surface Reconstruction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Poisson Surface Reconstruction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>"
)
```

Poisson surface reconstruction retrieves a closed, watertight surface mesh
from a point cloud equipped with oriented normals.

It operates in an [implicit space](https://en.wikipedia.org/wiki/Implicit_surface):
it solves a Poisson equation for the indicator function whose gradient best
matches the input normals, then extracts the zero level-set as a triangle mesh.

```{r setup}
library(vespa)
hand_url <- "https://gitlab.kitware.com/vtk/meshing/vespa/-/raw/master/Data/Testing/hand.vtp?ref_type=heads"
hand_tmp <- tempfile(fileext = ".vtp")
download.file(hand_url, hand_tmp, quiet = TRUE)
```

## Preparation

The demo uses `hand.vtp` from the VESPA test suite — a dense triangulated
surface mesh of a human hand (61 082 vertices, 122 160 triangles). We
deliberately discard the face connectivity and retain only the vertex
positions, turning the mesh into a raw, unstructured point cloud. This lets
us run the full reconstruction pipeline: normal estimation with PCA, then
Poisson reconstruction. The original mesh will serve as ground-truth for the
visual comparison at the end.


```{r build-pointcloud, eval = can_run}
hand_mesh <- read_vtp(hand_tmp)
print(hand_mesh)
```

We then strip the face connectivity to obtain a raw, unstructured point cloud:

```{r strip-faces, eval = can_run}
hand_pts <- hand_mesh
hand_pts$it <- matrix(integer(0), 3L, 0L)
```

## Normal estimation

`poisson_reconstruction()` requires oriented normals. If your point cloud has
none (as is the case here, since we stripped the mesh connectivity and with
it any per-vertex normal information), use `pca_estimate_normals()`: it fits
a local plane to the `n_neighbors` nearest neighbours and derives a normal
from that plane. Setting `orient = TRUE` propagates a consistent orientation
across the cloud via a minimum spanning tree on the normal graph.

```{r pca-normals, eval = can_run}
hand_pts <- pca_estimate_normals(hand_pts, n_neighbors = 18L, orient = TRUE)
```

## Poisson reconstruction

The three key parameters control the density and quality of the output mesh:

| Parameter    | Role |
|---|---|
| `min_angle`  | Minimum triangle angle (degrees). Higher values give more regular triangles but may increase face count. |
| `max_size`   | Maximum triangle edge length relative to the local point spacing. Lower values produce finer meshes. |
| `distance`   | Maximum Hausdorff distance between the output mesh and the implicit surface. Smaller values follow the surface more faithfully. |

```{r reconstruct, eval = can_run}
hand_recon <- poisson_reconstruction(
  hand_pts,
  min_angle = 20,
  max_size  = 2,
  distance  = 0.375
)
print(hand_recon)
```

## Comparison

```{r compare, eval = can_run}
cat("Input point cloud :", ncol(hand_pts$vb), "points\n")
cat("Reconstructed mesh:", ncol(hand_recon$vb), "vertices,",
    ncol(hand_recon$it), "triangles\n")
```

The Poisson method tends to produce a smooth, slightly blurred reconstruction
because the implicit solve acts as a low-pass filter. Tightening `distance`
and `max_size` recovers finer surface details at the cost of a higher face
count.

```{r rgl-compare, eval = FALSE}
rgl::mfrow3d(1, 2)
rgl::points3d(t(hand_pts$vb[1:3, ]), size = 1, col = "steelblue")
rgl::title3d("Input point cloud")
rgl::next3d()
rgl::shade3d(hand_recon, col = "tomato")
rgl::title3d("Poisson reconstruction")
```
