---
title: "Signed Distance Function"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Signed Distance Function}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(vespa)
dragon_url <- "https://gitlab.kitware.com/vtk/meshing/vespa/-/raw/master/Data/Testing/dragon.vtp"
dragon_tmp <- tempfile(fileext = ".vtp")
download.file(dragon_url, dragon_tmp, quiet = TRUE)
```

Speaking of implicit surface, the Signed Distance Function transforms a close-meshed shape into an implicit surface. In VTK terms, you can transform a PolyData into an ImageData. 

Here, many operations or visualizations are simpler like convoluting your shape with a kernel, having a volumetric render of your object, resampling, etc. You can then reconstruct a mesh using the Contour filter. Choosing an isosurface value different from 0 allows you to grow or shrink the original mesh.

```{r, fig.show='hold', fig.cap="a) Initial mesh; b) Shape thickening (isovalue > 0); c) Shape shrinking (isovalue < 0)", out.width='33%', echo=TRUE}
library(rgl)

dragon <- read_vtp(dragon_tmp)

# a) Original mesh
open3d()
shade3d(dragon, color = "lightblue")
title3d(main = "a) Original mesh")

# Build the SDF volume — result is now an `sdf_volume` object
sdf <- signed_distance_function(dragon, base_resolution = 128L, padding = 10L)
sdf 

# Isovalues expressed in physical units (voxel size)
iso_thicken <-  3 * sdf$spacing[1]   # outside -> grows
iso_shrink  <- -3 * sdf$spacing[1]   # inside  -> shrinks

# b) Thickened mesh
mesh_b <- extract_isosurface(sdf, isovalue = iso_thicken)
open3d()
shade3d(mesh_b, color = "lightgreen")
title3d(main = sprintf("b) Thickened (iso = %.2f)", iso_thicken))

# c) Shrunk mesh
mesh_c <- extract_isosurface(sdf, isovalue = iso_shrink)
open3d()
shade3d(mesh_c, color = "salmon")
title3d(main = sprintf("c) Shrunk (iso = %.2f)", iso_shrink))
```